Quantum Mechanics Assignment Nine

### 2016-04-05 solid_state/latex/a1.tex
Add SS stuff.

\title{Quantum Mechanics Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem 2.17}
This problem asks us to take the following spin-1 state and figure out the probabilities of
measuring z- and x-component spins, and also to find their expectation value.\\

\begin{align*}
  \ket{\Psi} \doteq &\frac{1}{\sqrt{30}}
  \begin{pmatrix}
    1\\
    2\\
    5i\\
  \end{pmatrix}
\end{align*}

We use the normal $P_x = \abs{\braket{\Psi_x}{\Psi}}^2$ to find probabilities. It is simple to
calculate the braket since both states are in the $S_z$ representation.

\begin{align*}
  P_{1} = \abs{\braket{\Psi_{1}}{\Psi}}^2 &=
  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}1&0&0\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
  = \frac{1}{30}\\
  P_{0} = \abs{\braket{\Psi_{0}}{\Psi}}^2 &=
  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}0&1&0\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
  = \frac{4}{30}\\
  P_{-1} = \abs{\braket{\Psi_{-1}}{\Psi}}^2 &=
  \abs{\frac{1}{\sqrt{30}}\begin{pmatrix}0&0&1\end{pmatrix}\begin{pmatrix}1\\2\\5i\\\end{pmatrix}}^2
  = \frac{25}{30}\\
\end{align*}
\begin{align*}
  \expval{S_z}{\Psi} &= \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}\begin{pmatrix}
      \hslash&0&0\\0&0&0\\0&0&-\hslash\\\end{pmatrix}\begin{pmatrix}1\\2\\5i\end{pmatrix}
        = \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}
          \begin{pmatrix}\hslash\\0\\5\hslash i\end{pmatrix}
            = \frac{1}{30}\left(1 + -25\right) = \frac{-4}{5}\\
\end{align*}

We then want to compute the expectation value of the spin in the x-direction. We use the
$S_z$ representation of $S_x$:

\begin{align*}
  \expval{S_x}{\Psi} &= \frac{1}{30}\begin{pmatrix}1&2&-5i\end{pmatrix}\frac{\hslash}{\sqrt{2}}
    \begin{pmatrix}0&1&0\\1&0&1\\0&1&0\\\end{pmatrix}\begin{pmatrix}1\\2\\5i\end{pmatrix}
        = \frac{1}{30\sqrt{2}}\begin{pmatrix}1&2&-5i\end{pmatrix}
          \begin{pmatrix}2\\1+5i\\2\end{pmatrix}\\
            = \frac{4}{30\sqrt{2}}\\
\end{align*}

\section{Problem 5.11}
This problem asks us to find the probabilities to find $n=2$ Hydrogen electrons in the classically
forbidden region.\\

This problem deals with an infinite square well wave which suddenly has its right potential wall
move from $L$ to $3L$. We are asked to find the probability of finding the wave in the ground or
first excited state. This problem is basically asking us to find the $c_1$ and $c_2$ coefficients
of the old wave function in the new well, and from them find probabilities.

\begin{align*}
  \Psi(x) &= \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)\\
  \ket{E_1} &= \sqrt{\frac{2}{3L}}\sin\left(\frac{\pi x}{3L}\right)\\
  \ket{E_2} &= \sqrt{\frac{2}{3L}}\sin\left(\frac{2\pi x}{3L}\right)\\
\end{align*}

We wish to find the $c_1$ and $c_2$ coefficients for this system.

\begin{align*}
  c_1 &= \braket{E_1}{\Psi} = \int_0^{3L} \sqrt{\frac{2}{3L}}\sin\left(\frac{\pi x}{3L}\right)*
  \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)dx\\
  &= \sqrt{\frac{4}{3L^3}}\int_0^{3L}
  \sin\left(\frac{\pi x}{3L}\right)\sin\left(\frac{\pi x}{L}\right)\\
  &= \sqrt{\frac{2}{3L^3}}\left(\int_0^{3L} \cos\left(\frac{2\pi x}{3L}\right)dx
  - \int_0^{3L} \cos\left(\frac{4\pi x}{3L}\right)dx\right)\\
  &= 0\\
\end{align*}

\begin{align*}
  c_2 &= \braket{E_2}{\Psi} = \int_0^{3L} \sqrt{\frac{2}{3L}}\sin\left(\frac{2\pi x}{3L}\right)*
  \sqrt{\frac{2}{L}}\sin\left(\frac{\pi x}{L}\right)dx\\
  &= \sqrt{\frac{4}{3L^3}}\int_0^{3L}
  \sin\left(\frac{2\pi x}{3L}\right)\sin\left(\frac{\pi x}{L}\right)\\
  &= \sqrt{\frac{2}{3L^3}}\left(\int_0^{3L} \cos\left(\frac{\pi x}{3L}\right)dx
  - \int_0^{3L} \cos\left(\frac{5\pi x}{3L}\right)dx\right)\\
  &= 0\\
\end{align*}

From this we can see that the probability of measuring the wave in the new tripled well to be in
$\ket{E_1}$ or $\ket{E_2}$ is zero, since $P_n = \abs{c_n}^2$. This result is to be expected, since
the wave should be completely in the second excited state, $\ket{E_3}$. This is because extending
the $\Psi(x)$ wave function from $0$ to $3L$ results in a two-node wave function identical to the
$\ket{E_3}$ state.\\

\section{8.7}
We begin with the potential and find the r-values at which $E=V$. We then calculate
the probability using $P = \int_r^\infty \psi^* \psi dr$.

\begin{align*}
  V(r) &= \frac{-e^2}{4\pi\epsilon_0r}\\
\end{align*}

We first solve for the forbidden distance:

\begin{align*}
  V(r) &= E_2\\
  -\frac{Z\alpha\hslash c}{r} &= -\frac{1}{2n^2}\alpha^2m_ec^2\\
  \frac{-1.44 \mbox{eV nm}}{r} &= -3.4 \mbox{eV}\\
  r &= 2.36 \mbox{nm}
\end{align*}

Next we find the probability of our particles being further than this:

\begin{align*}
  P_{200} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
  \frac{r^2}{32\pi a_0^3}\left(2-\frac{r}{a_0}\right)^2e^{\frac{-r}{a_0}}
  \sin(\theta)drd\phi d\theta = 2.09*10^{-14}\\
  P_{21-1} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
  \frac{r^4}{64\pi a_0^5}e^{\frac{-r}{a_0}}\sin^3(\theta)
  drd\phi d\theta = 7.62*10^{-15}\\
  P_{210} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
  \frac{r^4}{32\pi a_0^5}e^{\frac{-r}{a_0}}\cos^2(\theta)
  \sin(\theta)drd\phi d\theta = 7.62*10^{-15}\\
  P_{211} &= \int_{2.36}^\infty \int_0^{2\pi} \int_0^\pi
  \frac{r^4}{64\pi a_0^5}e^{\frac{-r}{a_0}}\sin^3(\theta)
  drd\phi d\theta = 7.62*10^{-15}\\
\end{align*}

We note that it is much more likely for the s-shell electron to be found in the forbidden (far
from atom) zone than the p-shell electrons. This is indicative of the shape of the s vs p shells
and how spread-out they are from the nucleus. We also note that, though we started off with 3
different wave functions for the $\ell = 1$ states, we get the same probabilities in the end.
This is to be expected, since the different $m_\ell$-numbered states are just rotations of each
other around the atom's axis, their shapes are the same.\\

I was initially confused as to why the units of the wave function were in $m^{-3/2}$, but then
recalled that complex conjugating the wave function and integrating over three dimensions would
cancel out these units.\\

>>>>

\title{Solid State Assignment One}
\author{Elliott Capek}
\titlespacing{\section}{0pt}{0pt}{5pt} % Spacing around titles

@@

\section{Problem 1: D functions as linear combinations}
For each of the five d-orbitals wavefunctions, create a 3D plot of the angular wave and express
the angular wave as a linear combination of spherical harmonics, like so:

\begin{align*}
  d_{x^2-y^2} &= a_{00}Y_{00} + a_{1-1}Y_{1-1} + a_{10}Y_{10} + a_{11}Y_{11} + a_{20}Y_{20} + ...\\
\end{align*}

\subsection{d

### 2016-04-05 solid_state/latex/a4.tex
Add SS stuff.

\begin{align*}
  H' &= \vec{\mu} \cdot \vec{B_2}\\
  &= \frac{g e}{2m} \hat{\mu} \cdot B_2 \hat{x}\\
  &= \frac{\hbar\omega_2}{2} S_x
  \hspace{2cm}\omega_2 = \frac{geB}{2m}\\
  &=
  \begin{pmatrix}
    0 & \frac{\hbar\omega_2}{2} & 0\\
    \frac{\hbar\omega_2}{2} & 0 & \frac{\hbar\omega_2}{2}\\
    0 & \frac{\hbar\omega_2}{2} & 0\\
  \end{pmatrix}
\end{align*}

@@

\begin{align*}
  E_1^{(2)} &=
  \frac{|H'_{1,2}|^2}{E_2 - E_1} +
  \frac{|H'_{1,3}|^2}{E_{3} - E_{1}}\\
  &= \frac{\hbar^2\omega_2^2}{4\hbar\omega_0}\\
  &= \frac{\hbar\omega_2^2}{4\omega_0}\\
\end{align*}

@@

\begin{align*}
  E_3^{(2)} &=
  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
  \frac{|H'_{3,2}|^2}{E_2 - E_3}\\
  &= -\frac{\hbar\omega_2^2}{4\omega_0}\\
\end{align*}

@@

@@

\begin{align*}
  E_3^{(2)} &=
  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
  \frac{|H'_{3,2}|^2}{E_2 - E_3} +
  \frac{|H'_{3,4}|^2}{E_4 - E_3}\\
  &= \frac{4\epsilon^2}{-2V_0} + \frac{\epsilon^2}{-4V_0}\\
  &= \frac{-9\epsilon^2}{4V_0}
\end{align*}

\begin{align*}
  E_4^{(2)} &=
  \frac{|H'_{4,1}|^2}{E_1 - E_4} +
  \frac{|H'_{4,2}|^2}{E_2 - E_4} +
  \frac{|H'_{4,3}|^2}{E_3 - E_4}\\
  &= \frac{-\epsilon^2}{2V_0}
\end{align*}

@@

>>>>

\begin{align*}
  H' &= -\vec{\mu} \cdot \vec{B_2}\\
  &= -\frac{g e}{2m} \hat{\mu} \cdot B_2 \hat{x}\\
  &= -\frac{\hbar\omega_2}{2} S_x
  \hspace{2cm}\omega_2 = \frac{geB}{2m}\\
  &=
  -\begin{pmatrix}
    0 & \frac{\hbar\omega_2}{\sqrt{2}} & 0\\
    \frac{\hbar\omega_2}{\sqrt{2}} & 0 & \frac{\hbar\omega_2}{\sqrt{2}}\\
    0 & \frac{\hbar\omega_2}{\sqrt{2}} & 0\\
  \end{pmatrix}
\end{align*}

@@

\begin{align*}
  E_1^{(2)} &=
  \frac{|H'_{1,2}|^2}{E_2 - E_1} +
  \frac{|H'_{1,3}|^2}{E_{3} - E_{1}}\\
  &= \frac{\hbar^2\omega_2^2}{4\hbar\omega_0}\\
  &= \frac{\hbar\omega_2^2}{4\omega_0}\\
  &= \frac{geB_2^2\hbar}{4mB_0}
\end{align*}

@@

\begin{align*}
  E_3^{(2)} &=
  \frac{|H'_{3,1}|^2}{E_1 - E_3} +
  \frac{|H'_{3,2}|^2}{E_2 - E_3}\\
  &= -\frac{\hbar\omega_2^2}{4\omega_0}\\
  &= \frac{-geB_2^2\hbar}{4mB_0}
\end{align*}

@@

\textbf{d.}
Here we must use wave function notation since the perturbed Hamiltonian is in the
spacial representation:

\begin{align*}
  E_1^{(1)} &= \bra{1^0}H'\ket{1^0}\\
  &= \frac{2V_0}{L\epsilon}
  \int_{\frac{L}{2}-\frac{\epsilon L}{2}}^{\frac{L}{2}+\frac{\epsilon L}{2}}
  \sin^2\left(\frac{\pi x}{L}\right)dx\\
  &= V_0 + \frac{2V_0\sin(\pi\epsilon)}{\pi\epsilon}
  \hspace{2cm}\mbox{Wolfram Alpha}\\
\end{align*}

\textbf{e.}
The limit of our energy correction as $\epsilon\rightarrow\infty$ is just $2V_0$,
as expected, since this case is identical to the delta function case.\\

@@

\begin{align*}
  E_3^{(2)} &=
  \frac{|H'_{3,1}|^2}{E_3 - E_1} +
  \frac{|H'_{3,2}|^2}{E_3 - E_2} +
  \frac{|H'_{3,4}|^2}{E_3 - E_4}\\
  &= \frac{4\epsilon^2}{2V_0} - \frac{\epsilon^2}{2V_0}\\
  &= \frac{3\epsilon^2}{2V_0}
\end{align*}

\begin{align*}
  E_4^{(2)} &=
  \frac{|H'_{4,1}|^2}{E_4 - E_1} +
  \frac{|H'_{4,2}|^2}{E_4 - E_2} +
  \frac{|H'_{4,3}|^2}{E_4 - E_3}\\
  &= \frac{\epsilon^2}{2V_0}
\end{align*}

@@

This means that the final corrected energies are:

\begin{align*}
  E_1 &= E_2 = 3 \pm \epsilon\\
\end{align*}

### 2016-04-05 solid_state/latex/a9.tex
Add SS stuff.

>>>>

\section{Problem three: p-state transitions}
Here we use McIntyre (14.84) to find which quantum states can be reached from the p-state.
From 14.84 we grab the following statement:\\

\begin{align*}
  \sqrt{\frac{3(2\ell_i+1)}{4\pi(2\ell_f+1)}}\bra{\ell_i,1,m_i,m}\ket{\ell_f,m_f}
  \bra{\ell_i,1,0,0}\ket{\ell_f,0}\\
\end{align*}

We then plug in values for $\ell_i$ and find the CB coefficients that lead to nonzero
values for this expression.\\

By looking at the CB table, we can see that for the $m_1=0,m_2=0$ RHS braket, the only
possible coupled basis J values are 0,2. This means that $\ell_f=0,2$, so by this we
can see that $\ell_i\neq\ell_f$. This is the only selection rule imposed by the RHS
braket, so we ignore it and look instead at the left braket:\\

\begin{align*}
  \bra{1,1,1,1}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,2}\\
  \bra{1,1,1,0}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,1}\\
  \bra{1,1,1,-1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
  \bra{1,1,0,1}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,1}\\
  \bra{1,1,0,0}\ket{\ell_f,m_f}  \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
  \bra{1,1,0,-1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,-1}\\
  \bra{1,1,-1,1}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,0},\braket{0,0}\\
  \bra{1,1,-1,0}\ket{\ell_f,m_f} \rightarrow\bra{\ell_f,m_f}=\braket{2,-1}\\
  \bra{1,1,-1,-1}\ket{\ell_f,m_f}\rightarrow\bra{\ell_f,m_f}=\braket{2,-2}\\
\end{align*}

From this we can see that the allowed transitions from the p-state are either to the s state
or any of the d-state orbitals.\\

This was a weird problem. It is odd that you can treat the final electron state as
represented in the coupled basis and the original state in the uncoupled basis and get
meaningful results. This is really cool, but an odd usage of the CB table.\\
